### 2024-12-11 paper/paper.tex
paper

%    Copyright (c)  2024  João Augusto Costa Branco Marado Torres.
%    Permission is granted to copy, distribute and/or modify this document
%    under the terms of the GNU Free Documentation License, Version 1.3
%    or any later version published by the Free Software Foundation;
%    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
%    A copy of the license is included in the section entitled "GNU
%    Free Documentation License".

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Define Article %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Using Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{empheq}
\usepackage{mdframed}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{color}
\usepackage{psfrag}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{bm}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\selectlanguage{english}
\usepackage{blindtext}
\usepackage{csquotes}
\usepackage{hyperref}
\hypersetup{colorlinks,
           citecolor=black,
           filecolor=black,
           linkcolor=black,
           urlcolor=black,
           bookmarksopen=true,
           pdftex}
\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\addbibresource{references.bib}
\usepackage{bookmark}
\usepackage{enumitem}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot[C]{\copyrightnotice}
\setlength{\headheight}{14.5pt}
%\addtolength{\topmargin}{-2.5pt}
\usepackage{mathtools}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Other Settings

\pagenumbering{arabic}
\hfuzz = .6pt % avoid black boxes

%%%%%%%%%%%%%%%%%%%%%%%%%% Page Setting %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\geometry{a4paper}

%%%%%%%%%%%%%%%%%%%%%%%%%% Define some useful colors %%%%%%%%%%%%%%%%%%%%%%%%%%
\definecolor{ocre}{RGB}{243,102,25}
\definecolor{mygray}{RGB}{243,243,244}
\definecolor{deepGreen}{RGB}{26,111,0}
\definecolor{shallowGreen}{RGB}{235,255,255}
\definecolor{deepBlue}{RGB}{61,124,222}
\definecolor{shallowBlue}{RGB}{235,249,255}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Define an orangebox command %%%%%%%%%%%%%%%%%%%%%%%%
\newcommand\orangebox[1]{\fcolorbox{ocre}{mygray}{\hspace{1em}#1\hspace{1em}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\copyrightnotice}{
    Copyright \copyright{}  2024  João Augusto Costa Branco Marado Torres.
}
\newcommand{\licensenotice}{
    \copyrightnotice
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.3
    or any later version published by the Free Software Foundation;
    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled ``GNU
    Free Documentation License''.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% English Environments %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newtheoremstyle{mytheoremstyle}{3pt}{3pt}{\normalfont}{0cm}{\rmfamily\bfseries}{}{1em}{{\color{black}\thmname{#1}~\thmnumber{#2}}\thmnote{\,--\,#3}}
\newtheoremstyle{myproblemstyle}{3pt}{3pt}{\normalfont}{0cm}{\rmfamily\bfseries}{}{1em}{{\color{black}\thmname{#1}~\thmnumber{#2}}\thmnote{\,--\,#3}}
\theoremstyle{mytheoremstyle}
\newmdtheoremenv[linewidth=1pt,backgroundcolor=shallowGreen,linecolor=deepGreen,leftmargin=0pt,innerleftmargin=20pt,innerrightmargin=20pt,]{theorem}{Theorem}[section]
\theoremstyle{mytheoremstyle}
\newmdtheoremenv[linewidth=1pt,backgroundcolor=shallowBlue,linecolor=deepBlue,leftmargin=0pt,innerleftmargin=20pt,innerrightmargin=20pt,]{definition}{Definition}[section]
\theoremstyle{myproblemstyle}
\newmdtheoremenv[linecolor=black,leftmargin=0pt,innerleftmargin=10pt,innerrightmargin=10pt,]{problem}{Problem}[section]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Plotting Settings %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepgfplotslibrary{colorbrewer}
\pgfplotsset{width=8cm,compat=1.9}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title & Author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{LANGuage IDentification}
\author{João Augusto Costa Branco Marado Torres}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
    \maketitle

    \section*{License}
    \bigskip
    \begin{quote}
        \licensenotice
    \end{quote}
    \bigskip

    \tableofcontents
    %\clearpage

    %\addcontentsline{toc}{chapter}{Foreword}
    %{\huge {\bf Foreword}}

    % \Blindtext

    \section{Phrases to Numeric Representation}

    I was really confused about this, but it really is not that complicated.

    Before starting to write any code, I wanted to know how can I transform a
    phrase like the one you are reading right now into \texttt{1}s and
    \texttt{0}s because the computer isn't smart enough to know English.
    So I went to my friend and
    \href{https://chatgpt.com/share/675976d6-b2c8-8002-964c-a3fff698bcc0}{asked
    it some questions}.
    It thought me various ways of achieving what I wanted, and between those
    options I picked the easiest to understand/implement.

    \subsection{Bag of Words}

    The idea it's to build a \textbf{vocabulary} which is a list of unique
    words.

    Everyone (including us) has its own vocabulary composed by words of the
    languages we speak.

    Our AI model needs a vocabulary too so he can understand some words. Those
    words might not exist in our vocabulary.

    Let's say \hyperref[tab:rainbow_model_vocab]{our model's vocabulary} are
    the words in English for the 7 colors in the rainbow.

    Each word in the vocabulary will have a numerical identifier.

    \begin{table}
        \caption{Rainbow model vocabulary}\label{tab:rainbow_model_vocab}
        \begin{center}
            \begin{tabular}[c]{l|l}
                \hline
                \multicolumn{1}{c|}{\textbf{ID}} &
                \multicolumn{1}{c}{\textbf{Word}} \\
                \hline
                0 & red \\
                1 & orange \\
                2 & yellow \\
                3 & green \\
                4 & cyan \\
                5 & blue \\
                6 & violet \\

                \hline
            \end{tabular}
        \end{center}
    \end{table}

    It makes sense for us for that identifiers to start at \texttt{0} and
    increment by \texttt{1} for each word.

    Now our model can represent our phrases into something it understands.

    \begin{quote}
        How?
    \end{quote}

    The question you just asked is represented as a zero column matrix with 7
    rows.

    The model will try to find words he knows from a phrase and represent it as
    a column matrix with rows the same as the amount of words in the vocabulary.
    Each entry represents a word in the vocabulary, here is why it's a good
    idea to identify the words as I said. The value of each entry can be a
    simple boolean that represent if a word is present in the phrase or not, or
    the amount of times the word appears in the phrase. The value really
    represents how important is the that word in the phrase.

    The phrase ``How?'' does not contain any words know by the model.

    The phrase ``The colors in the \textit{Guiné-Bissau} flag are red, yellow,
    green and black in the star.'' would be represented by the matrix
    $\begin{bsmallmatrix}1&0&1&1&0&0&0\end{bsmallmatrix}^T$ and the phrase
    ``What came first, the orange fruit or the orange color?'' could be
    represented by $\begin{bsmallmatrix}0&2&0&0&0&0&0\end{bsmallmatrix}^T$ if
    we count the amount of occurrences of the word in the phrase.

    The way I implemented the BoW was by reading a file and collect every
    single word from it separated by white spaces, dashes, punctuation,
    parenthesis or quotation marks and then spit to \texttt{stdout} a formated
    vocabulary, so you can pipe it later or write it to a file. Words are
    case-sensitive, and it's possible to create the vocabulary in alphabetic
    order.

    My implementation can be found in \href{run:../src/commands/vocab/bow.zig}{\texttt{/src/commands/vocab/bow.zig}}.

    During the conversation, ``tokenization'' was mentioned, and it might be
    cool for you to take a look at it.

    \subsection{Other ways}

    \begin{itemize}
        \item TF-IDF;
        \item One-Hot Encoding of Words;
        \item Embedding-like Approach;
        \item Sub-word tokenization.
    \end{itemize}

    \medskip

    \printbibliography[
    heading=bibintoc
    ]

\end{document}

### 2024-12-11 paper/build.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+# This is part of the LANGuage IDentification.
+# Copyright (C) 2024
+#   João Augusto Costa Branco Marado Torres
+# See the file ./paper.tex for copying conditions.
+
+set -xe
+
+pdflatex paper
+# bibtex8 paper
+biber paper
+pdflatex paper
+pdflatex paper

### 2024-12-12 paper/paper.tex
paper: neuron

@@
 \hypersetup{colorlinks,
-           citecolor=black,
-           filecolor=black,
-           linkcolor=black,
-           urlcolor=black,
-           bookmarksopen=true,
-           pdftex}
+           %citecolor=black,
+           %filecolor=black,
+           %linkcolor=black,
+           %urlcolor=black,
+           bookmarksopen=true}
 \usepackage[
@@
 \usepackage{mathtools}
+\usepackage{tikz}
+\usetikzlibrary{positioning,fit,calc}
+\usepackage{float}
+\usepackage{wrapfig}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+    \listoffigures
+    %\listoftables
+
     \section{Phrases to Numeric Representation}
@@
 
-    Let's say \hyperref[tab:rainbow_model_vocab]{our model's vocabulary} are
+    Let's say {our model's vocabulary}~\ref{tab:rainbow_model_vocab} are
     the words in English for the 7 colors in the rainbow.
@@
     The model will try to find words he knows from a phrase and represent it as
-    a column matrix with rows the same as the amount of words in the vocabulary.
-    Each entry represents a word in the vocabulary, here is why it's a good
-    idea to identify the words as I said. The value of each entry can be a
-    simple boolean that represent if a word is present in the phrase or not, or
-    the amount of times the word appears in the phrase. The value really
-    represents how important is the that word in the phrase.
+    a column matrix with rows the same as the amount of words in the
+    vocabulary. Each entry represents a word in the vocabulary, here is why
+    it's a good idea to identify the words as I said. The value of each entry
+    can be a simple boolean that represent if a word is present in the phrase
+    or not, or the amount of times the word appears in the phrase. The value
+    really represents how important is the that word in the phrase.
 
@@
     single word from it separated by white spaces, dashes, punctuation,
-    parenthesis or quotation marks and then spit to \texttt{stdout} a formated
+    parenthesis or quotation marks and then spit to \texttt{stdout} a formatted
     vocabulary, so you can pipe it later or write it to a file. Words are
@@
 
-    My implementation can be found in \href{run:../src/commands/vocab/bow.zig}{\texttt{/src/commands/vocab/bow.zig}}.
+    My implementation can be found in
+    \href{run:../src/commands/vocab/bow.zig}{\texttt{/src/commands/vocab/bow.zig}}.
 
@@
 
+    \section{Neuron}
+
+    As explained at the start of chapter 3.1 of \cite{bishop}, the base for the
+    simplest supervised linear regression models will look like
+    \eqref{eq:linear_model}
+    \begin{equation}
+        \begin{split}
+            y(\mathbf{x}, \mathbf{w}) & = (1)b + w_{0}x_{0} + w_{1}x_{1} +
+            \ldots + w_{D}x_{D} = \\ & = (1)b +
+            \displaystyle\sum_{j=0}^{D}\left(w_{j}x_{j}\right)
+        \end{split}
+        \label{eq:linear_model}
+    \end{equation}
+    where $ \mathbf{x} =
+    \begin{bsmallmatrix}x_{0}&\cdots&x_{D}\end{bsmallmatrix}^{T} $, a linear
+    function of the parameters $ \mathbf{w} =
+    \begin{bsmallmatrix}w_{0}&\cdots&w_{D}\end{bsmallmatrix} $ and bias $b$. We
+    can also make the bias part of the $ \mathbf{w} $ and at the same position
+    in $ \mathbf{x} $ put the value 1 and have the equation like
+    \eqref{eq:linear_model_matrix}
+    \begin{equation}
+        y(\mathbf{x}, \mathbf{w}) = \mathbf{x}\cdot\mathbf{w}
+        \label{eq:linear_model_matrix}
+    \end{equation}
+    a simple matrix multiplication. Both matrixes have the same amount of
+    elements $ D $.
+
+    There is also the notion of applying what is called as \textbf{basis
+    function} to the $ \mathbf{x} $ parameters before everything. So there will
+    be a matrix of basis functions $ \mathbf{\phi} =
+    \begin{bsmallmatrix}\phi_{0}&\cdots&\phi_{D}\end{bsmallmatrix}^{T} $ where
+    $ b = x_{i} $ and $ \phi_{i}(b) = 1 $.
+
+    If that's the case the function will look like this
+    \eqref{eq:linear_model_basis_f}
+    \begin{equation}
+        \begin{split}
+            y(\mathbf{x}, \mathbf{w}) & =
+            \displaystyle\sum_{j=0}^{D}\left(w_{j}\phi_{j}(x_{j})\right) = \\
+            & = \mathbf{\phi}\left(\mathbf{x}\right)\cdot\mathbf{w}
+        \end{split}
+        \label{eq:linear_model_basis_f}
+    \end{equation}
+
+    The functions can be seen as a way to pre-process and extract the important
+    parts of the input This list can have nonlinear functions making $
+    y(\mathbf{x}, \mathbf{w}) $ nonlinear too.
+
+    In case we are working with a classification model, we might want to
+    introduce an \textbf{activation function} $ f\left(\cdot\right) $ that
+    after all the computations in $ y(\mathbf{x}, \mathbf{w}) $
+    \eqref{eq:linear_model_basis_f} transforms that result into a value
+    corresponding to being the probability of $ \mathbf{x} $ being of a certain
+    class.
+
+    And this will be what I define as an artificial neuron \eqref{eq:neuron}.
+    You can see a visual representation of it on page~\pageref{fig:neuron}
+
+    \begin{equation}
+        y(\mathbf{x}, \mathbf{w}) = f \left( \mathbf{\phi} \left( \mathbf{x}
+        \right) \cdot\mathbf{w} \right)
+        \label{eq:neuron}
+    \end{equation}
+
+    \begin{figure}
+        \begin{center}
+            \begin{tikzpicture}[
+                    neuron/.style={circle, draw=black!90, very thick},
+                    var/.style={}, io/.style={circle, draw=black!90,
+                    fill=black!10, very thick, minimum size=7mm},
+                ]
+                \node (neuron_logic) {$ y\left(\mathbf{x}, \mathbf{w}\right)
+                $};
+                \node (weights) [above=of neuron_logic] {$ \mathbf{w} $};
+                \node[text opacity=0] (invisible) [below=of neuron_logic] {neuron};
+                \node[neuron, fit={(weights)(neuron_logic)(invisible)}] (neuron) {};
+                % \node[var] (input) [left=of neuron] {$ \mathbf{x} = \begin{bmatrix}x_{0}&x_{1}&\vdot&x_{D}\end{bmatrix} $};
+                \node[var] (input) [left=of neuron] {$ \mathbf{x} $};
+                \node[var] (bias) [above=of input] {$ b = x_{0} $};
+                \node[var] (output) [right=of neuron] {$ y $};
+                \node[io] (basis) at (neuron.west) {$
+                \mathbf{\phi}\left(\mathbf{x}\right) $};
+                \node[io] (activation) at (neuron.east) {$ f\left(y\right) $};
+
+                \draw[->] (bias.south) -- (input.north);
+                \draw[->] (input.east) -- (basis.west);
+                \draw[->] (basis.east) -- (neuron_logic.west);
+                \draw[->] (neuron_logic.east) -- (activation.west);
+                \draw[->] (activation.east) -- (output.west);
+                \draw[->] (weights.south) -- (neuron_logic.north);
+            \end{tikzpicture}
+        \end{center}
+        \caption{A visual representation of a neuron.}\label{fig:neuron}
+    \end{figure}
+
+    Because I couldn't find a simple way of doing anonymous functions in my
+    language of choice, \href{run:../src/utils/neuron.zig}{my implementation}
+    does not allow specifying basis functions nor activation functions so every
+    basis function can be thought as being the identify function $
+    \mathbf{\phi}(x) = x $ and the activation function can be toggled between
+    the identify and sigmoid \eqref{eq:sigmoid}.
+
+    \begin{equation}
+        \sigma\left(x\right) = \frac{1}{1+e^{-x}}
+        \label{eq:sigmoid}
+    \end{equation}
+
+    %\section{Feed-Forward Neural Network}
+
     \medskip

### 2024-12-13 paper/paper.tex
neural network and forward propagation

@@
 %    Free Documentation License".
+%
+%    You should have received a copy of the GNU Free Documentation License
+%    along with this program.  If not, see <https://www.gnu.org/licenses/>
 
@@
     If that's the case the function will look like this
-    \eqref{eq:linear_model_basis_f}
+    \eqref{eq:linear_model_basis_f}.
     \begin{equation}
@@
     And this will be what I define as an artificial neuron \eqref{eq:neuron}.
-    You can see a visual representation of it on page~\pageref{fig:neuron}
+    You can see a visual representation of it on page~\pageref{fig:neuron}.
 
@@
     Because I couldn't find a simple way of doing anonymous functions in my
-    language of choice, \href{run:../src/utils/neuron.zig}{my implementation}
+    language of choice, \href{run:../src/utils/neural.zig}{my implementation}
     does not allow specifying basis functions nor activation functions so every
@@
 
-    %\section{Feed-Forward Neural Network}
+    \section{Feed-Forward Neural Network}
+
+    The idea now, as explained in chapter 5.1 of \cite{bishop}, is to train
+    neurons so that their weight values are good.
+
+    Because a neuron is really just a function, a neuron can be a basis
+    function of another neuron, starting a network.
+
+    %%% create visual representation of the above
+
+    A network is composed of layers, each with its own number of neurons.
+
+    %%% visual representation of a layer
+
+    Each of neuron will create a value called \textbf{activation} represented
+    by $ a_{j} $ \eqref{eq:activation} where $ j $ is the neuron number (we can
+    think of a layer as a column matrix of neurons), $ \mathbf{X} $ is the
+    function input or the activations from the layer before after they passed
+    through an activation function and $ D $ is the amount of row in $
+    \mathbf{x} $ $ - 1 $.
+
+    Let's define a way to represent a matrix row:
+    $$
+        A = \begin{bmatrix}a_{0}\\\vdots\\a_{N}\end{bmatrix} \in \mathbb{R}^{N
+            \times M}, a_{i} \in \mathbb{R}^{M}, i \in \left\{ x \in
+            \mathbb{N}: x \le N \right\}
+    $$
+    $$
+        A_{i *} = a_{i}
+    $$
+
+    \begin{equation}
+        \begin{split}
+            a_{j} & = f \left( \displaystyle\sum_{i=1}^{D} \left( w_{ji}x_{i}
+            \right) + (1)b_{j} \right) \\ & = f \left( \mathbf{x} \cdot W_{j *}
+            \right)
+        \end{split}
+        \label{eq:activation}
+    \end{equation}
+
+    The last layer will be the one giving the outputs $ \mathbf{y} $.
+
+    If a neural network has $ n $ layers, the mathematical representation for
+    the output $ k $ will be something like \eqref{eq:neural_network}. $
+    W^{(n)} $ are the weights of that layer. For this to work, because it's
+    just matrix multiplications, we have to be aware of the dimensions of the
+    matrixes. This is called \textbf{forward propagation}.
+
+    \begin{equation}
+        Y_{0} = \textbf{x}
+    \end{equation}
+    \begin{equation}
+        Y_{n} = f_{n} \left( Y_{n - 1}^{T} \cdot w^{(n)} \right), n > 0
+    \end{equation}
+
+    \begin{equation}
+        \begin{split}
+            y_{k}(\mathbf{x}, \mathbf{w}) & = f_{n} \left( \displaystyle\sum_{i_{n}=0} w_{k i_{n}}^{(n)} \cdots f_{2} \left( \displaystyle\sum_{i_{2}=0} w_{i_{3}i_{2}}^{(2)} f_{1} \left( \displaystyle\sum_{i_{1}=0} w_{i_{2}i_{1}}^{(1)}x_{i_{1}} \right) \right) \cdots \right) = \\
+            & = f_{n} \left( \cdots f_{2} \left( f_{1} \left( \mathbf{x} \cdot w^{(1)} \right)^{T} \cdot w^{(2)} \right)^{T} \cdots \cdot w^{(n)} \right) = \\
+            & = Y_{n}
+        \end{split}
+        \label{eq:neural_network}
+    \end{equation}
+
+    %%% full conectivity not mandatory
+
+    %%% visual representation of a neural network
+
+    Layers and neural network representation can be found
+    \href{run:../src/utils/neural.zig}{\texttt{/src/utils/neural.zig}}.
+
+    %%% create x^2 sin(x) |x|
 

### 2024-12-17 paper/paper.tex
basic understanding of back propagation

@@
 
-    \section*{License}
+    \section{License}
     \bigskip
@@
         \begin{split}
-            y_{k}(\mathbf{x}, \mathbf{w}) & = f_{n} \left( \displaystyle\sum_{i_{n}=0} w_{k i_{n}}^{(n)} \cdots f_{2} \left( \displaystyle\sum_{i_{2}=0} w_{i_{3}i_{2}}^{(2)} f_{1} \left( \displaystyle\sum_{i_{1}=0} w_{i_{2}i_{1}}^{(1)}x_{i_{1}} \right) \right) \cdots \right) = \\
-            & = f_{n} \left( \cdots f_{2} \left( f_{1} \left( \mathbf{x} \cdot w^{(1)} \right)^{T} \cdot w^{(2)} \right)^{T} \cdots \cdot w^{(n)} \right) = \\
+            y_{k}(\mathbf{x}, \mathbf{w}) & = f_{n} \left( \displaystyle\sum_{i_{n}=0} w_{k i_{n}}^{(n)} \hdots f_{2} \left( \displaystyle\sum_{i_{2}=0} w_{i_{3}i_{2}}^{(2)} f_{1} \left( \displaystyle\sum_{i_{1}=0} w_{i_{2}i_{1}}^{(1)}x_{i_{1}} \right) \right) \hdots \right) = \\
+            & = f_{n} \left( \hdots f_{2} \left( f_{1} \left( \mathbf{x} \cdot w^{(1)} \right)^{T} \cdot w^{(2)} \right)^{T} \hdots \cdot w^{(n)} \right) = \\
             & = Y_{n}
@@
 
+    \section{Learning}
+
+    Now we can feed our network with some data, and it will give us a response
+    (one that we might not want). The network will probably give you some
+    nonsense output and that's probably because the weights in each neuron are
+    also nonsense.
+
+    Our goal now it's to try and find the optimal weights for the network.
+
+    We will evaluate how bad the network is behaving and after that tell it
+    what it should change to do a better job.
+
+    \subsection{Error function}
+
+    First we need a way of measuring how dumb the network is. Error functions
+    measures the difference between the network output and the expected output.
+
+    There are a lot of functions that we can choose to calculate that.
+
+    We will use this $ E \left( \mathbf{w} \right) $ error function
+    \eqref{eq:error_function} that is a sum of the squares of those differences
+    in outputs.
+
+    \begin{equation}
+        E_{n} \left( \mathbf{w} \right) = \left| \mathbf{y}(\mathbf{x}_{n}, \mathbf{w}) - \mathbf{t}_{n} \right|^{2}
+    \end{equation}
+
+    \begin{equation}
+        E \left( \mathbf{w} \right) = \frac{1}{2} \displaystyle\sum_{n=1}^{N} E_{n}(\mathbf{w})
+        \label{eq:error_function}
+    \end{equation}
+
+    We have the input training matrix $ \left\{ \mathbf{x}_{n} \right\} $, $ N $
+    is the input size. Also, we have $ \left\{ \mathbf{t}_{n} \right\} $ with the
+    expected outputs for each input.
+
+    Our goal is to minimize the error function \eqref{eq:error_function}
+    because the closer to 0 that the error function return gets, means the
+    differences between our neural network output and the expected output are
+    less big.
+
+    \subsection{Parameter Optimization}
+
+    The way to minimize a function is by calculating its derivative on a point,
+    so we can get the instantaneous rate of change at that point, that is, how
+    much is the function growing/decreasing on that point. But that only works
+    for unary function. Our neural network is a function with an unlimited
+    (limited by the hardware which is running on) number of parameters (the
+    weights). Because of that we instead use \textbf{partial derivatives}, and
+    we calculate it for each one of the neural network weights (and biases if
+    they are separated from the weights). Actually, we are calculation what's
+    called the \textbf{gradient}, which is a vector of which one of those
+    partial derivatives. That vector's direction is towards where that function
+    grows (greatest rate of increase), we want to go the other way around.
+
+    Our goal is to reach a minimum with $ \nabla E \left( \mathbf{w} \right) =
+    \vec{0} $ and while we don't reach it, we keep changing the weights based
+    on $ -\nabla E \left( \mathbf{w} \right) $. One of the problems that might
+    arise is that $ \nabla E \left( \mathbf{w} \right) = \vec{0} $ does not
+    mean that we found the perfect weight values, because a function can have
+    multiple minimums and with the amount of parameters that a neural network
+    has, it surely does, but also, it does not mean for sure we meet a minimum
+    and instead meet a maximum of saddle point. Also targeting non error can be
+    impossible, so we might instead target a maximum error that we allow the
+    neural network to have.
+
+    It also probably hard as hell to find the perfect weights only using math;
+    I wouldn't get involved in that. At least it would take some time. So
+    normally we just give each weight a random value, and we have the initial
+    weight vector $ \mathbf{w}^{\left( 0 \right)} $. From that, we iteratively
+    update the weights based on a weight vector update $ \Delta
+    \mathbf{w}^{\left( \tau \right)} $ as shown in \eqref{eq:weight_updates}.
+
+    \begin{equation}
+        \mathbf{w}^{\left( \tau + 1 \right)} = \mathbf{w}^{\left( \tau \right)} + \Delta \mathbf{w}^{\left( \tau \right)}
+        \label{eq:weight_updates}
+    \end{equation}
+
+    In our case we are going to do \textbf{gradient descent optimization} so
+    from \eqref{eq:weight_updates} we get
+    \eqref{eq:gradient_descent_optimization}.
+
+    \begin{equation}
+        \begin{split}
+            \mathbf{w}^{\left( \tau + 1 \right)} & = \mathbf{w}^{\left( \tau \right)} + \Delta \mathbf{w}^{\left( \tau \right)} = \\
+            & = \mathbf{w}^{\left( \tau \right)} + \left( \left( \mathbf{w}^{\left( \tau \right)} + \eta \nabla E \left( \mathbf{w}^{\left( \tau \right)} \right) \right) - \mathbf{w}^{\left( \tau \right)} \right) = \\
+            & = \mathbf{w}^{\left( \tau \right)} + \eta \nabla E \left( \mathbf{w}^{\left( \tau \right)} \right)
+        \end{split}
+        \label{eq:gradient_descent_optimization}
+    \end{equation}
+
+    The $ \eta > 0 $ and it's called \textbf{learning rate}, so we can define
+    how fast we want the optimization to be based on the gradient.
+
+    While implementing this, I had some problems.
+
+    Sometimes the $ E \left( \mathbf{w} \right) $ was minimizing and from out of nowhere it
+    would explode into big numbers. But it seems there are some techniques that
+    seem to maybe solve that problem such as \textbf{regularization} which is
+    mentioned in page 30 \cite{bishop}. It can also be because of the way I
+    calculated the derivatives: by using the limit definition. You can define
+    an $ \epsilon $ on your neural network, and it will calculate the
+    derivatives using it like this: $ \lim_{\epsilon\to\infty} \frac{f(x +
+    \epsilon) - f(x)}{\epsilon} $. It would also depend on the input, and the
+    initial weights.
+
+    During implementation, I also thought that this system can cause a lot of
+    problems in multilayer neural networks, because changing the weights of one
+    neuron will affect the behavior of every neuron forward because they depend
+    on its activation, and we are changing every weight at the same time.
+
+    But I guess that's why \textbf{error backpropagation} is preferred over
+    this method.
+
+    \subsection{Error Backpropagation}
+
+    Maybe because of what I mentioned above, error backpropagation exist, where
+    the gradient is calculated from the last layer of the neural network (the
+    output layer) to the first.
+
+    This should work for any feed-forward neural network with any activation
+    functions and error function.
+
+    We have a general error function which is the sum of the errors of each
+    input data.
+
+    \begin{equation}
+        E \left( \mathbf{w} \right) = \displaystyle\sum_{n=1}^{N} E_{n}(\mathbf{w})
+    \end{equation}
+
+    We want to compute $ \nabla E_{n} \left( \mathbf{w} \right) $.
+
+    So we start by feeding forward where we calculate all the activations for
+    every input $ \mathbf{x}_{i} $. In all the equations forward $ \left( \tau
+    \right) $ represents the layer.
+
+    Remember from the equation \eqref{eq:linear_model}. Here we have
+
+    \newcommand{\nlayer}{\left( \tau \right)}
+    \newcommand{\nplayer}[1]{\left( \tau - {#1} \right)}
+    \newcommand{\npblayer}[1]{\left( \tau + {#1} \right)}
+    \begin{equation}
+        a_{k}^{\nlayer} = \displaystyle\sum_{i} \left( w_{ki}^{\nlayer} z_{i}^{\nplayer{1}} \right)
+    \end{equation}
+
+    \begin{equation}
+        z_{i}^{\left( \tau \right)} = f^{\left( \tau \right)} \left( a_{i}^{\left( \tau \right)} \right)
+    \end{equation}
+
+    \newcommand{\pd}[2]{\frac{\partial {#1}}{\partial {#2}}}
+    \newcommand{\ajl}{a_{j^{\nlayer}}^{\nlayer}}
+    \newcommand{\ajpl}[1]{a_{j^{\npblayer{#1}}}^{\npblayer{#1}}}
+    Now as we did before, we want to calculate the gradient, so we need to
+    calculate the partial derivative $ \pd{E_{n}}{w_{ji}^{\nlayer}} $ for every
+    weight.
+
+    We can use the derivatives chain rule to write the partial derivative
+    because the error function needs the activation after the weight to be
+    dependent on that weight.
+
+    \begin{equation}
+        \pd{E_{n}}{w_{ji}^{\nlayer}} & = \pd{E_{n}}{a_{j}^{\nlayer}} \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}}
+    \end{equation}
+
+    And calculate those two separately.
+
+    \begin{equation}
+        \begin{split}
+            \pd{E_{n}}{a_{j}^{\nlayer}} & = \pd{}{a_{j}^{\nlayer}} \left( a_{j}^{\left( \tau \right)} - t_{j} \right)^{2} = \\
+            & = 2 \left( a_{j}^{\left( \tau \right)} - t_{j} \right) \pd{}{a_{j}^{\nlayer}} \left( a_{j}^{\left( \tau \right)} - t_{j} \right) = \\
+            & = 2 \left( a_{j}^{\left( \tau \right)} - t_{j} \right) \left( 1 \right) = \\
+            & = 2 \left( a_{j}^{\left( \tau \right)} - t_{j} \right) = \\
+            & = \delta_{j}
+        \end{split}
+    \end{equation}
+    \begin{equation}
+        \begin{split}
+            \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}} & = \pd{}{w_{ji}^{\nlayer}} \displaystyle\sum_{i} \left( w_{ji}^{\nlayer} z_{i}^{\nplayer{1}} \right) = \\
+            & = \pd{}{w_{ji}^{\nlayer}} \left( w_{ji}^{\nlayer} z_{i}^{\nplayer{1}} + \displaystyle\sum_{k \neq i} \left( w_{jk}^{\nlayer} z_{k}^{\nplayer{1}} \right) \right) = \\
+            & = \left( 1 \right) z_{i}^{\nplayer{1}} + \displaystyle\sum_{k \neq i} 0 = \\
+            & = z_{i}^{\nplayer{1}}
+        \end{split}
+    \end{equation}
+
+    \begin{equation}
+        \pd{E_{n}}{w_{ji}} = \delta_{j}^{\nlayer} z_{i}^{\nplayer{1}}
+    \end{equation}
+
+    We call $ \delta_{j}^{\nlayer} $ the errors. Because we already calculated
+    every $ z_{i}^{\nplayer{1}} $ while forward propagating, we really just
+    need to calculate the errors.
+
+    The $ \delta_{j}^{\nlayer} $ we calculated is for the output layer. For the
+    layer before that, we can also use the chain rule.
+
+    \begin{equation}
+        \begin{split}
+            \delta_{j}^{\nplayer{1}} & = \pd{E_{n}}{a_{j}^{\nplayer{1}}} \\
+            & = \displaystyle\sum_{k} \left( \pd{E_{n}}{a_{k}^{\nlayer}} \pd{a_{k}^{\nlayer}}{a_{j}^{\nplayer{1}}} \right) = \\
+            & = \displaystyle\sum_{k} \left( \delta_{k}^{\nlayer} \pd{}{a_{j}^{\nplayer{1}}} \left( w_{kj}^{\nlayer} f^{\nplayer{1}} \left( a_{j}^{\nplayer{1}} \right) + \displaystyle\sum_{i \neq j} \left( w_{ki}^{\nlayer} z_{ki}^{\nplayer{1}} \right) \right) \right) = \\
+            & = \displaystyle\sum_{k} \left( \delta_{k}^{\nlayer} w_{ki}^{\nlayer} f^{\prime \nplayer{1}} \left( a_{i}^{\nplayer{1}} \right)  \right) = \\
+            & = f^{\prime \nplayer{1}} \left( a_{j}^{\nplayer{1}} \right) \displaystyle\sum_{k} \left( w_{kj}^{\nlayer} \delta_{k}^{\nlayer} \right)
+        \end{split}
+    \end{equation}
+
+    \begin{equation}
+        \begin{split}
+            \pd{E_{n}}{w_{ji}^{\nlayer}} & = \pd{E_{n}}{a_{j}^{\nlayer}} \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}} = \\
+            & = \pd{E_{n}}{\ajpl{l}} \displaystyle\prod_{L = 1}^{l} \left( \pd{\ajpl{l - \left( L - 1 \right)}}{\ajpl{l - L}} \right) \pd{\ajl}{w_{ji}^{\nplayer{l}}}
+        \end{split}
+    \end{equation}
+
+    After calculating all errors from the output layer, you are able to
+    calculate from the other layers, one at a time, backwards, and you'll end
+    up with the derivatives for the gradient.
+
+    We may also accumulate gradients to apply it to the weights all at once.
+
+    \begin{equation}
+        \pd{E}{w_{ji}^{\nlayer}} & = \displaystyle\sum_{n} \pd{E_{n}}{w_{ji}^{\nlayer}}
+    \end{equation}
+
     \medskip

### 2024-12-18 paper/paper.tex
start reviewing paper

@@
 %    along with this program.  If not, see <https://www.gnu.org/licenses/>
-
+%
+%    <https://github.com/Marado-Programmer/langid>
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%% Define Article %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 style=alphabetic,
-sorting=ynt
+sorting=ynt,
+defernumbers=true
 ]{biblatex}
+\DeclareBibliographyCategory{cited}
+\AtEveryCitekey{\addtocategory{cited}{\thefield{entrykey}}}
 \addbibresource{references.bib}
@@
 
+\nocite{*}
+
 \begin{document}
@@
 
-    \section{License}
+    \begin{abstract}
+        This paper goes through the basics of supervised machine learning while
+        simultaneously I implement with code what I'm learning without the use
+        of libraries that hide what's really happening underneath and don't
+        allow us to really understand and learn.
+
+        I'm writing this paper because of school. But still, this is a topic I
+        would like to really understand.
+
+        Because now AI in general has become main stream it feels like that
+        what already seemed like a black box for me, it has become even more
+        abstracted away, so the average person can make use of these
+        technologies.
+
+        During my last internship (January to March 2023) there was a time I
+        had to do some audio recognition stuff that I remember used
+        \href{https://www.tensorflow.org/js}{tensorflow, the ML library for
+        JavaScript}, under the hood.
+
+        Good for me that there was an interface for the things I needed to
+        implement, but if the guy that was giving me tasks had an idea in which
+        those interfaces couldn't help, I would not know what to do.
+
+        I know what I learned doing this project it's just the surface, but
+        it's this foundation that I need to start learning more complex ideas.
+
+        The project repository is available at: \\
+        \url{https://github.com/Marado-Programmer/langid}
+    \end{abstract}
+
+    \section*{License}
+
     \bigskip
@@
     %\listoftables
+
+    \section{Introduction}
+
+    The initial goal was to create a model that would be trained on unlabeled
+    phrases in different languages, and after that you would be able to prompt
+    any phrase, and then it would tell which language was it written in.
+
+    Turns out I didn't even get into unsupervised learning on this project. So
+    the data is labeled.
+
+    Still, I learned a lot and this paper will share some of the things I
+    learned.
+
+    Most things I learned/relearned (some things was talked about in class) by
+    reading \cite{bishop} but when it comes to the backpropagation algorithm,
+    if it wasn't for the playlist \cite{3b1b} I probably would still be stuck
+    there.
 
@@
     \begin{equation}
-        \pd{E_{n}}{w_{ji}^{\nlayer}} & = \pd{E_{n}}{a_{j}^{\nlayer}} \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}}
+        \pd{E_{n}}{w_{ji}^{\nlayer}} = \pd{E_{n}}{a_{j}^{\nlayer}} \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}}
     \end{equation}
@@
     \begin{equation}
-        \pd{E}{w_{ji}^{\nlayer}} & = \displaystyle\sum_{n} \pd{E_{n}}{w_{ji}^{\nlayer}}
+        \pd{E}{w_{ji}^{\nlayer}} = \displaystyle\sum_{n} \pd{E_{n}}{w_{ji}^{\nlayer}}
     \end{equation}
+
+    \section{Some Observations}
+
+    I think the number of layers and total of neurons should grow when the
+    input size grows too. If the input is two numbers and I want their
+    multiplication as the output, 1 single neuron layer will not be able to do
+    a great job.
+
+    I also observed that the learning rate should decrease with the increase of
+    layers/neurons or else the weights can become big numbers.
 
@@
     \printbibliography[
-    heading=bibintoc
+    heading=bibintoc,
+    category=cited
+    ]
+
+    \printbibliography[
+    heading=bibintoc,
+    notcategory=cited,
+    title={Other Resources}
     ]

### 2024-12-18 paper/paper.tex
paper reviewed

@@
 \usepackage{wrapfig}
+\usepackage{gnuplot-lua-tikz}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
         The project repository is available at: \\
-        \url{https://github.com/Marado-Programmer/langid}
+        \url{https://github.com/Marado-Programmer/langid} \cite{repo}
     \end{abstract}
@@
 
-    The way I implemented the BoW was by reading a file and collect every
-    single word from it separated by white spaces, dashes, punctuation,
+    The way I implemented the Bag of Words was by reading a file and collect
+    every single word from it separated by white spaces, dashes, punctuation,
     parenthesis or quotation marks and then spit to \texttt{stdout} a formatted
@@
     language of choice, \href{run:../src/utils/neural.zig}{my implementation}
-    does not allow specifying basis functions nor activation functions so every
+    does not allow specifying basis functions so every
     basis function can be thought as being the identify function $
-    \mathbf{\phi}(x) = x $ and the activation function can be toggled between
-    the identify and sigmoid \eqref{eq:sigmoid}.
+    \mathbf{\phi}(x) = x $, but the activation functions can be specified via
+    an enumeration value which gives a set of options like \eqref{eq:sigmoid}
+    for example.
 
@@
     Because a neuron is really just a function, a neuron can be a basis
-    function of another neuron, starting a network.
+    function of another neuron, starting a network~\ref{fig:neurons_connecting}.
 
-    %%% create visual representation of the above
+    \begin{figure}
+        \begin{center}
+            \begin{tikzpicture}[
+                    neuron/.style={circle, draw=black!90, very thick},
+                    var/.style={}, io/.style={circle, draw=black!90,
+                    fill=black!10, very thick, minimum size=7mm},
+                ]
+                \node[neuron] (neuron_1) {1};
+                \node[neuron] (neuron_2) [right=of neuron_1] {2};
+                \node[var] (input) [left=of neuron_1] {$ \mathbf{x} $};
+                \node[var] (output) [right=of neuron_2] {$ y $};
 
-    A network is composed of layers, each with its own number of neurons.
+                \draw[->] (input.east) -- (neuron_1.west);
+                \draw[->] (neuron_1.east) -- (neuron_2.west);
+                \draw[->] (neuron_2.east) -- (output.west);
+            \end{tikzpicture}
+        \end{center}
+        \caption{Neuron 1 working as a basis function for neuron 2.}\label{fig:neurons_connecting}
+    \end{figure}
 
-    %%% visual representation of a layer
+    A network~\ref{fig:network} is composed of layers, each with its own number of neurons.
 
-    Each of neuron will create a value called \textbf{activation} represented
-    by $ a_{j} $ \eqref{eq:activation} where $ j $ is the neuron number (we can
-    think of a layer as a column matrix of neurons), $ \mathbf{X} $ is the
-    function input or the activations from the layer before after they passed
-    through an activation function and $ D $ is the amount of row in $
-    \mathbf{x} $ $ - 1 $.
+    Each one of the neurons will create a value called \textbf{activation}
+    represented by $ a_{j} $ \eqref{eq:activation} where $ j $ is the neuron
+    number (we can think of a layer as a column matrix of neurons), $
+    \mathbf{X} $ is the function input or the activations from the layer before
+    after they passed through an activation function and $ D $ is the amount of
+    row in $ \mathbf{x} $ $ - 1 $.
 
@@
     \begin{equation}
-        Y_{0} = \textbf{x}
+        Y_{0} = \mathbf{x}
     \end{equation}
@@
 
-    %%% visual representation of a neural network
+    \begin{figure}
+        \begin{center}
+            \begin{tikzpicture}[
+                    neuron/.style={circle, draw=black!90, very thick},
+                    var/.style={}, io/.style={circle, draw=black!90,
+                    fill=black!10, very thick, minimum size=7mm},
+                ]
+                \node[neuron] (neuron_1_1) {};
+                \node[neuron] (neuron_1_2) [below=of neuron_1_1] {};
+                \node[neuron] (neuron_1_3) [below=of neuron_1_2] {};
+                \node[neuron] (neuron_1_4) [below=of neuron_1_3] {};
+                \node[neuron] (neuron_1_5) [below=of neuron_1_4] {};
+                \node[neuron] (neuron_2_1) [right=2cm of neuron_1_2] {};
+                \node[neuron] (neuron_2_2) [right=2cm of neuron_1_4] {};
+                \node[var] (input_m) [left=2cm of neuron_1_3] {$ \vdots $};
+                \node[var] (input_i) [above=of input_m] {$ \mathbf{x}_{1} $};
+                \node[var] (input_f) [below=of input_m] {$ \mathbf{x}_{D} $};
+                \node[var] (output) [below right=of neuron_2_1] {$ y $};
+                \node[var] (layer1) [above=of neuron_1_1] {1\textsuperscript{st} layer};
+                \node[var] (layer2) [above=of neuron_2_1] {2\textsuperscript{nd} layer};
 
-    Layers and neural network representation can be found
+                \draw[->] (input_i.east) -- (neuron_1_1.west);
+                \draw[->] (input_i.east) -- (neuron_1_2.west);
+                \draw[->] (input_i.east) -- (neuron_1_3.west);
+                \draw[->] (input_i.east) -- (neuron_1_4.west);
+                \draw[->] (input_i.east) -- (neuron_1_5.west);
+                \draw[->] (input_f.east) -- (neuron_1_1.west);
+                \draw[->] (input_f.east) -- (neuron_1_2.west);
+                \draw[->] (input_f.east) -- (neuron_1_3.west);
+                \draw[->] (input_f.east) -- (neuron_1_4.west);
+                \draw[->] (input_f.east) -- (neuron_1_5.west);
+                \draw[->] (neuron_1_1.east) -- (neuron_2_1.west);
+                \draw[->] (neuron_1_1.east) -- (neuron_2_2.west);
+                \draw[->] (neuron_1_2.east) -- (neuron_2_1.west);
+                \draw[->] (neuron_1_2.east) -- (neuron_2_2.west);
+                \draw[->] (neuron_1_3.east) -- (neuron_2_1.west);
+                \draw[->] (neuron_1_3.east) -- (neuron_2_2.west);
+                \draw[->] (neuron_1_4.east) -- (neuron_2_1.west);
+                \draw[->] (neuron_1_4.east) -- (neuron_2_2.west);
+                \draw[->] (neuron_1_5.east) -- (neuron_2_1.west);
+                \draw[->] (neuron_1_5.east) -- (neuron_2_2.west);
+                \draw[->] (neuron_2_1.east) -- (output.west);
+                \draw[->] (neuron_2_2.east) -- (output.west);
+            \end{tikzpicture}
+        \end{center}
+        \caption{Two layer neural network, first layer has 5 neurons and the
+        second one 2.}\label{fig:network}
+    \end{figure}
+
+    Layers and neural network representation can be found \\
     \href{run:../src/utils/neural.zig}{\texttt{/src/utils/neural.zig}}.
-
-    %%% create x^2 sin(x) |x|
 
@@
             \mathbf{w}^{\left( \tau + 1 \right)} & = \mathbf{w}^{\left( \tau \right)} + \Delta \mathbf{w}^{\left( \tau \right)} = \\
-            & = \mathbf{w}^{\left( \tau \right)} + \left( \left( \mathbf{w}^{\left( \tau \right)} + \eta \nabla E \left( \mathbf{w}^{\left( \tau \right)} \right) \right) - \mathbf{w}^{\left( \tau \right)} \right) = \\
-            & = \mathbf{w}^{\left( \tau \right)} + \eta \nabla E \left( \mathbf{w}^{\left( \tau \right)} \right)
+            & = \mathbf{w}^{\left( \tau \right)} + \left( \left( \mathbf{w}^{\left( \tau \right)} - \eta \nabla E \left( \mathbf{w}^{\left( \tau \right)} \right) \right) - \mathbf{w}^{\left( \tau \right)} \right) = \\
+            & = \mathbf{w}^{\left( \tau \right)} - \eta \nabla E \left( \mathbf{w}^{\left( \tau \right)} \right)
         \end{split}
@@
     We have a general error function which is the sum of the errors of each
-    input data.
+    input data \eqref{eq:general_error_function}.
 
@@
         E \left( \mathbf{w} \right) = \displaystyle\sum_{n=1}^{N} E_{n}(\mathbf{w})
+        \label{eq:general_error_function}
     \end{equation}
@@
     every input $ \mathbf{x}_{i} $. In all the equations forward $ \left( \tau
-    \right) $ represents the layer.
+    \right) $ represents the layer $ \tau $ being the last layer.
 
-    Remember from the equation \eqref{eq:linear_model}. Here we have
+    Remember from the equation \eqref{eq:neuron}. Here we have
 
@@
     Now as we did before, we want to calculate the gradient, so we need to
-    calculate the partial derivative $ \pd{E_{n}}{w_{ji}^{\nlayer}} $ for every
-    weight.
+    calculate the partial derivative $ \pd{E_{n}}{w_{ji}^{\nlayer}} $
+    \eqref{eq:weight_partial_derivative} for every weight.
 
@@
         \pd{E_{n}}{w_{ji}^{\nlayer}} = \pd{E_{n}}{a_{j}^{\nlayer}} \pd{a_{j}^{\nlayer}}{w_{ji}^{\nlayer}}
+        \label{eq:weight_partial_derivative}
     \end{equation}
@@
         \end{split}
+        \label{eq:errors}
     \end{equation}
@@
         \pd{E_{n}}{w_{ji}} = \delta_{j}^{\nlayer} z_{i}^{\nplayer{1}}
+        \label{eq:weight_partial_derivative_simplified}
     \end{equation}
 
-    We call $ \delta_{j}^{\nlayer} $ the errors. Because we already calculated
-    every $ z_{i}^{\nplayer{1}} $ while forward propagating, we really just
-    need to calculate the errors.
+    We can rewrite \eqref{eq:weight_partial_derivative} as
+    \eqref{eq:weight_partial_derivative_simplified}.
+
+    We call $ \delta_{j}^{\nlayer} $ \eqref{eq:errors} the errors. Because we
+    already calculated every $ z_{i}^{\nplayer{1}} $ while forward propagating,
+    we really just need to calculate the errors.
 
     The $ \delta_{j}^{\nlayer} $ we calculated is for the output layer. For the
-    layer before that, we can also use the chain rule.
+    layer before that, we can also use the chain rule and calculate the other
+    errors \eqref{eq:other_weight_partial_derivative}.
 
@@
         \end{split}
+        \label{eq:other_weight_partial_derivative}
     \end{equation}
@@
 
-    We may also accumulate gradients to apply it to the weights all at once.
+    We may also accumulate gradients to apply it to the weights all at once
+    \eqref{eq:batch}.
 
@@
         \pd{E}{w_{ji}^{\nlayer}} = \displaystyle\sum_{n} \pd{E_{n}}{w_{ji}^{\nlayer}}
+        \label{eq:batch}
     \end{equation}
@@
     \section{Some Observations}
+
+    % \begin{figure}
+    %     \begin{center}
+    %         %\includegraphics[width=0.95\textwidth]{./res/double.jpeg}
+    %         \input{./res/doubles.tex}
+    %     \end{center}
+    %     \caption{Change of error function on neural network to compute $ y = 2x
+    %     $. The function y is the value the network is returning for one of the
+    %     input training data.}\label{fig:doubles}
+    % \end{figure}
+    \begin{wrapfigure}{O}{0.5\textwidth}
+        \begin{center}
+            \input{./res/abs.tex}
+        \end{center}
+        \caption{Change of error function on neural network to compute $ y =
+        \left| x \right| $. The function y is the value the network is
+        returning for one of the input training data.}\label{fig:abs}
+    \end{wrapfigure}
+    \medskip
+    % \begin{wrapfigure}{L}{0.48\textwidth}
+    %     \begin{center}
+    %         \input{./res/sin.tex}
+    %     \end{center}
+    %     \caption{Change of error function on neural network to compute $ y =
+    %     \sin \left( x \right) $. The function y is the value the network is
+    %     returning for one of the input training data.}\label{fig:sin}
+    % \end{wrapfigure}
+    % \begin{wrapfigure}{R}{0.48\textwidth}
+    %     \begin{center}
+    %         \input{./res/linear.tex}
+    %     \end{center}
+    %     \caption{Change of error function on neural network to compute $ y = ax
+    %     + b $. The function y is the value the network is returning for one of
+    %     the input training data.}\label{fig:linear}
+    % \end{wrapfigure}
+    % \begin{wrapfigure}{L}{0.48\textwidth}
+    %     \begin{center}
+    %         \input{./res/parabola.tex}
+    %     \end{center}
+    %     \caption{Change of error function on neural network to compute $ y =
+    %     ax^{2} + bx + c $. The function y is the value the network is returning
+    %     for one of the input training data.}\label{fig:parabola}
+    % \end{wrapfigure}
 
@@
     layers/neurons or else the weights can become big numbers.
+
+    There are so many think we can tweak. The learning rate, amount of layers
+    and neurons, the training data, activation functions, initial weights and
+    biases, error function. You really just need to stir the pile until the
+    results look okay \cite{xkcd1838}.
 

### 2024-12-18 paper/create_plots.sh

@@ -1,0 +1,46 @@
+#!/bin/sh
+# This is part of the LANGuage IDentification.
+# Copyright (C) 2024
+#   João Augusto Costa Branco Marado Torres
+# See the file ./paper.tex for copying conditions.
+
+set -xe
+
+mkdir -p ./res
+
+tmp=$(mktemp)
+
+../zig-out/bin/idlang eg doubles 20 \
+    --layers=1-1 \
+    --learning-rate=5e-3 \
+    --learning-amount=300 > "$tmp"
+gnuplot -c ./create_plots.gnuplot \
+    "$tmp" ./res/doubles.tex "\$ y = 2x \$"
+
+../zig-out/bin/idlang eg abs 20 \
+    --layers=4-2-1 \
+    --learning-rate=5e-3 \
+    --learning-amount=500 > "$tmp"
+gnuplot -c ./create_plots.gnuplot \
+    "$tmp" ./res/abs.tex "\$ y = \left| x \right| \$"
+
+../zig-out/bin/idlang eg sin 12 \
+    --layers=6-4-2-1 \
+    --learning-rate=1e-4 \
+    --learning-amount=1000 > "$tmp"
+gnuplot -c ./create_plots.gnuplot \
+    "$tmp" ./res/sin.tex "\$ y = \sin \left( x \right) \$"
+
+../zig-out/bin/idlang eg linear 60 \
+    --layers=4-2-1 \
+    --learning-rate=5e-4 \
+    --learning-amount=1000 > "$tmp"
+gnuplot -c ./create_plots.gnuplot \
+    "$tmp" ./res/linear.tex "\$ y = ax + b \$"
+
+../zig-out/bin/idlang eg parabola 80 \
+    --layers=8-4-4-1 \
+    --learning-rate=5e-6 \
+    --learning-amount=1500 > "$tmp"
+gnuplot -c ./create_plots.gnuplot \
+    "$tmp" ./res/parabola.tex "\$ y = ax^2 + bx + c \$"
